Kalman tracking and estimation of continuous gravitational waves with a pulsar timing array

### 2023-08-07 docs/manuscript/paper.tex
canonical example with a fixed h prior

\subsection{Model selection}\label{sec:model_selection}
The evidence integral,$\mathcal{Z}$, returned by nested sampling is the probability of the data $\boldsymbol{Y}$ given a particular model $\mathcal{M}_i$. This enables us to compare competing models via a Bayes factor,
\begin{equation}
	\beta = \frac{\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1)}{\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_0)} \ .
\end{equation}
Throughout this work we take $\mathcal{M}_1$ to be the complete model defined in Section \ref{sec:model}. $\mathcal{M}_0$ is our null model that assumes there is no GW in the data. This is equivalent to setting $g^{(n)}(t)=1$ in Equation \ref{eq:g_func}, \ref{eq:g_func_trig}. The Bayes factors we quote in this work therefore quantify whether the data supports evidence for a GW compared to there being no GW signal present.

@@

\begin{figure*}%
	\centering
	\subfloat[\centering $f_{\rm em}^{(n)}(t_1)$]{{\includegraphics[width=0.3 \textwidth]{images/example_residuals_plot} }}%
	\qquad
	\subfloat[\centering $\dot{f}_{\rm em}^{(n)}(t_1)$]{{\includegraphics[width=0.3 \textwidth]{images/example_residuals_plot} }}%
	\qquad
	\subfloat[\centering $\sigma^{(n)}$]{{\includegraphics[width=0.3\textwidth]{images/example_residuals_plot} }}%

	\caption{Actual (left panels) and synthetic (right panels) phase residuals for 3 NANOGrav pulsars. The actual residuals are obtained via the NANOGrav 12.5 year data release \citep{nanograv_narrowband_DR}. The synthetic residuals are generated by numerically solving Eq. \ref{eq:spinevol} with $\gamma^{(n)} = 10^{-13}$ and $\sigma^{(n)}$ inferred from Eqs. \ref{eq:sigmap}, \ref{eq:sigmap_f}. For all pulsars the two solutions are qualitatively similar, with comparable magnitudes in the residuals. \textcolor{red}{TK: these plots need updating, but need to discuss with Andres re use of baboo. Maybe just one is sufficient just as an example.}}
	\label{fig:qualitative_compare}
\end{figure*}

@@

\subsubsection{Detection}

We frame the detection problem in terms of model selection as described in Section \ref{sec:model_selection}

The variance of the measurement noise $\Sigma$ can be estimated approximately from the root-mean-square amplitude of phase residuals and we do not treat it as a parameter to be inferred in this work.

@@

@@

>>>>

\subsection{Model selection}\label{sec:model_selection}
The evidence integral,$\mathcal{Z}$, returned by nested sampling is the probability of the data $\boldsymbol{Y}$ given a particular model $\mathcal{M}_i$. This enables us to compare competing models via a Bayes factor,
\begin{equation}
	\beta = \frac{\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1)}{\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_0)} \ . \label{eq:bayes}
\end{equation}
Throughout this work we take $\mathcal{M}_1$ to be the complete model defined in Section \ref{sec:model}. $\mathcal{M}_0$ is our null model that assumes there is no GW in the data. This is equivalent to setting $g^{(n)}(t)=1$ in Equation \ref{eq:g_func}, \ref{eq:g_func_trig}. The Bayes factors we quote in this work therefore quantify whether the data supports evidence for a GW compared to there being no GW signal present.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/example_residuals_plot}
	\caption{Actual (left panels) and synthetic (right panels) phase residuals for 3 NANOGrav pulsars. The actual residuals are obtained via the NANOGrav 12.5 year data release \citep{nanograv_narrowband_DR}. The synthetic residuals are generated by numerically solving Eq. \ref{eq:spinevol} with $\gamma^{(n)} = 10^{-13}$ and $\sigma^{(n)}$ inferred from Eqs. \ref{eq:sigmap}, \ref{eq:sigmap_f}. For all pulsars the two solutions are qualitatively similar, with comparable magnitudes in the residuals. \textcolor{red}{TK: these plots need updating, but need to discuss with Andres re use of baboo. Maybe just one is sufficient just as an example.}}
	\label{fig:qualitative_compare}
\end{figure}

%
%\begin{figure*}%
%	\centering
%	\subfloat[\centering $f_{\rm em}^{(n)}(t_1)$]{{\includegraphics[width=0.3 \textwidth]{images/example_residuals_plot} }}%
%	\qquad
%	\subfloat[\centering $\dot{f}_{\rm em}^{(n)}(t_1)$]{{\includegraphics[width=0.3 \textwidth]{images/example_residuals_plot} }}%
%	\qquad
%	\subfloat[\centering $\sigma^{(n)}$]{{\includegraphics[width=0.3\textwidth]{images/example_residuals_plot} }}%
%	
%	\caption{Actual (left panels) and synthetic (right panels) phase residuals for 3 NANOGrav pulsars. The actual residuals are obtained via the NANOGrav 12.5 year data release \citep{nanograv_narrowband_DR}. The synthetic residuals are generated by numerically solving Eq. \ref{eq:spinevol} with $\gamma^{(n)} = 10^{-13}$ and $\sigma^{(n)}$ inferred from Eqs. \ref{eq:sigmap}, \ref{eq:sigmap_f}. For all pulsars the two solutions are qualitatively similar, with comparable magnitudes in the residuals. \textcolor{red}{TK: these plots need updating, but need to discuss with Andres re use of baboo. Maybe just one is sufficient just as an example.}}
%	\label{fig:qualitative_compare}
%\end{figure*}

@@

\subsubsection{Detection}
We frame the detection problem in terms of a Bayesian model selection procedure as described in Section \ref{sec:model_selection}.

The Bayes factor $\beta$ as defined in Eq. \ref{eq:bayes} is presented in Fig \ref{fig:bayes} for our representative example system, where we now vary the magnitude of the GW strain, $h_0$.

The noise is an identical realisation for each value of $h_0$

We take $\beta = 10$ as a tolerance cutoff above which we can accept $\mathcal{M}_1$ and claim a detection of the GW.

For this system, given this realisation of the noise, the minimal detectable strain is $\sim 3 \times 10^{-15}$

The small perturbations in the general log-linear $\beta - h_0$ trend arise due to noise in the nested sampling method. For these points the sampler did not converge as well for $\mathcal{M}_1$ and so the evidence is lower than would be expected. This is purely a noise artefact: running the nested sampler again and recreating the $\beta - h_0$ leads to either no perturbations, or perturbations at different points in the curve.

Below $\beta = 10$ this noise in the sampler starts to dominate and the hierarchical structure of the models $\mathcal{M}_0 , \mathcal{M}_1$ fails. That is, the requirement that $\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1) > \mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_0)$ no longer holds

and the $\beta - h_0$ curve becomes much less smooth

This is simply stating that at these small strains the system is completely noise dominated and the two models $\mathcal{M}_1, \mathcal{M}_0$ cannot be separated.

\begin{figure}
	\includegraphics[width=\columnwidth]{images/PaperBayesRatioPlotn1000_stacked}
	\caption{}
	\label{fig:bayes}
\end{figure}

@@

\subsection{Bias by dropping pulsar terms}

some example likelihood curves.

why are some affected but not others?

@@

comparison with alternative methods
